### 2020-09-24 Manuscript/What.lies.ahead.reviewed.tex
minor text fixes and editing.

Studies of the ecological effects of global change often focus on one or few species at a time. Consequently, we know relatively little about the changes underway at real-world scales of biological communities, which typically have hundreds or thousands of interacting species. Here, we use monthly samples of environmental DNA to survey over 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat. The result is a high-resolution picture of changes in ecological communities using a technique replicable across a wide variety of ecosystems. We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification. We find distinct communities in warmer and more acidified conditions, with overall reduced richness in diatom assemblages and increased richness in dinoflagellates. Individual taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful dinoflagellate \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.

@@

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time \cite{kelly2019understanding}, using pooled data from technical replicates and mean proportions across biological replicates. We used this index of abundance in the multivariate community analysis, and used binary (presence/absence) data to capture species-level responses to environmental conditions. {\color{red}An implementation in R of this index is provided in the Supplementary information.}

@@

Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community. The SIMPER procedure in \textit{vegan} revealed the taxa most contributing to between-cluster differences.

For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs. San Juan Island), as predictor variables. Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models. We calculated the probability of each community, given these predictors, using the r package \textit{nnet} \cite{nnet2002}.

@@

Given the sea-surface temperatures and pH values for 2017 (observed) and 2095 (estimated) and taxon-specific logistic regression models, we then evaluated the suitability of habitat for each taxon in the future scenario. For each point in the pH - temperature grid, we calculated a species' probability of presence as the mean of 100 independent draws of the posterior model response. For each point, the sum of mean probabilities across species provided richness estimations. The mean value across the 100 draws was the input for a Wilcoxon test of differences in species richness between the 2017 and 2095 scenarios in each region. We performed the Wilcoxon test globally (total species richness) and on each phylum independently. 
% Reviewer suggests that the paragraoh below belongs to the discussion
{\color{red} RYAN: Move paragraph below to discussion?}
Taxa surveyed are a function of our metabarcoding PCR primers \cite{leray_new_2013} and reflect the current status of genetic databases, rather than a complete sampling of the planktonic community; we view these results as a cross-section of common taxa useful for understanding the biological effects of ocean conditions.

@@

Metabarcoding analysis of eDNA samples generated more than ***50.8M sequences from 778 samples. These samples represented biological and technical replicates from 86 unique sampling events. After bioinformatics quality-control \st{(See Methods)} the dataset included $\sim${}***45M sequences, from ***4849 unique amplicon sequence variants (ASVs). Of these, ***1364 ASVs (22.6M reads) could be annotated to a taxonomic level of Family or lower. These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs. planktonic; see Supporting Information). Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 221). These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig. 1B.1) \cite{dethier_decadal_2008}.

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area ***(Hood Canal vs. San Juan Island; F-model = 1.6184; p $<$ 0.01). We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2; for clarity, results for Hood Canal shown; full analysis in the Supporting Information).

@@

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the community linked with colder water and higher-pH (community **2 in Fig. 2C) showed higher eDNA indexes of diatoms ***(\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp than that on the low pH spectrum. In that community (cluster 1) \textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} and the copepod \textit{Paracalanus} sp. C are more prevalent.  The community thriving in the high temperature range (community 3 in Fig. 2C) shows high values of \textit{Phaeocystis globosa},\textit{Ditylum brightwellii} among other species. Community 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 66\% of the time, an increase of 11\% compared to 2017(see Supplemental Material). On the other hand, conditions that make cluster 2 the most likely community drop from 9\% to 1\% of the time between 2017 and 2095.

% Under a common environmental condition, for example, in Hood Canal in summer, we expect such a planktonic community to go from occurring less than 1\% of the time in 2017 (0.001\% at T $=$ 15$\degree$C and pH $=$ 8.4; conditions more extreme than this occur ca. 50\% of the time) to occurring 37.5\% of the time under an equivalently-common condition in 2095 (84\% at T $=$ 22$\degree$C and pH $=$ 8.05).

@@

Diatoms (Bacillariophyta) show the steepest richness decline under future conditions (Figure 3 B.2); the probability of occurrence decreases markedly for diverse diatom taxa including both pennate (e.g., \textit{Nitzschia}) and centric (e.g., \textit{Coscinodiscus}, \textit{Stephanodiscus}) body forms. These declines in diatom richness were more accentuated at lower pH values and higher temperatures. Other taxa likely to find less-suitable habitat in the future include dinoflagellates \textit{Karlodinium} sp. and Gymnodiniaceae sp.

@@

 Changes in environmental conditions and associated shifts in planktonic communities will likely reshape ecosystems and food webs, although some environmental processes may be conserved even as the particular taxa change. A switch from a diatom-dominated ecosystem to one in which dinoflagellate blooms extend in space and time could provoke cascade effects \cite{xiao2018warming} including fish mortality, anoxia \cite{anderson2002harmful}, and {\color{red} carbon sinking dynamics \st{alter current bentho-pelagic coupling processes} \cite{bienfang1984sinking}}. Beyond the phylum-specific patterns, the increase in suitable habitat for harmful algae species will alone be an engine for ecosystem change \cite{trainer2019pelagic, wells2018harmful}.

@@

Our observations are strong evidence of the kinds of changes likely in future marine communities, and they offer testable predictions about the magnitude and direction of effects on focal species.

%%% Material y metodos

@@

 We are grateful to the UW Center for Environmental Genomics and Linda Park's lab at NOAA Fisheries for access to lab equipment and associated expertise and support. Thanks to Micah Horwith and the Washington State Department of Natural Resources for in-field support, to Terrie Klinger for feedback on an earlier draft, to Stephanie Moore for advice on phytoplankton generally, to Ana Ram\'{o}n-Laca for lab help, to Dan Drinan for field help, and to Alex Gagnon for access to carbonate-chemistry expertise and equipment.
 % Display the acknowledgments section

>>>>

Studies of the ecological effects of global change often focus on one or few species at a time. Consequently, we know relatively little about the changes underway at real-world scales of biological communities, which typically have hundreds or thousands of interacting species. Here, we use monthly samples of environmental DNA to survey 222 planktonic taxa along a gradient of temperature, salinity, dissolved oxygen, and carbonate chemistry in nearshore marine habitat. The result is a high-resolution picture of changes in ecological communities using a technique replicable across a wide variety of ecosystems. We estimate community-level differences associated with time, space and environmental variables, and use these results to forecast near-term community changes due to warming and ocean acidification. We find distinct communities in warmer and more acidified conditions, with overall reduced richness in diatom assemblages and increased richness in dinoflagellates. Individual taxa finding greater suitable habitat in near-future waters are more taxonomically varied and include the ubiquitous coccolithophore \textit{Emiliania huxleyi} and the harmful dinoflagellate \textit{Alexandrium sp.} These results suggest foundational changes for nearshore food webs under near-future conditions.

@@

By treating amplification efficiency as consistent within a given taxon, we created an index of abundance for each taxon across space and time (``eDNA index'' \cite{kelly2019understanding}), using pooled data from technical replicates and mean proportions across biological replicates. We used this index of abundance in the multivariate community analysis, and used binary (presence/absence) data to capture species-level responses to environmental conditions. In the Supplementary information, we provide a function in R to calculate this index.

@@

Independent of environmental parameters, we separately clustered samples by pairwise Bray-Curtis dissimilarities (k-means; N = 3) to identify groups of samples that were similar to one another with respect to biological community. The SIMPER procedure in \textit{vegan} revealed the taxa most strongly contributing to between-cluster differences.

For community-level projections, we coded community-cluster identity (Figure 2) as an unordered response variable in a multinomial logistic model, with temperature, pH, and area (Hood Canal vs. San Juan Island), as predictor variables. Salinity is predicted to remain largely unchanged in future scenarios \cite{khangaonkar2019salish}, and because salinity was correlated with temperature in our dataset, it was not an important predictor variable and we subsequently dropped it from our models. We calculated the probability of each community, given these predictors, using a multinomial logistic regression in the R package \textit{nnet} \cite{nnet2002}.

@@

Given the sea-surface temperatures and pH values for 2017 (observed) and 2095 (estimated) and taxon-specific logistic regression models, we then evaluated the suitability of habitat for each taxon in the future scenario. For each point in the pH - temperature grid, we calculated a species' probability of presence as the mean of 100 independent draws of the posterior model response. For each point, the sum of mean probabilities across species provided richness estimations. The mean value across the 100 draws was the input for a Wilcoxon test of differences in species richness between the 2017 and 2095 scenarios in each region. We performed the Wilcoxon test globally (total species richness) and on each phylum independently.

@@

Metabarcoding analysis of eDNA samples generated more than ***50.8M sequences from 778 samples. These samples represented biological and technical replicates from 86 unique sampling events. After bioinformatics quality-control the dataset included $\sim${}***45M sequences, from ***4849 unique amplicon sequence variants (ASVs). Of these, ***1364 ASVs (22.6M reads) could be annotated to a taxonomic level of Family or lower. These $\sim$ 500 taxa from 43 phyla were split according to their natural history and habitat (benthic vs. planktonic; see Supporting Information). Because we expect planktonic taxa to vary with water mass \cite{kelly2018tides} and therefore with bottle-sampled carbonate chemistry, here we focus on only the planktonic taxa (N = 221). These taxa showed a seasonal richness gradient between study areas, consistent with documented biodiversity clines in the area (Fig. 1B.1) \cite{dethier_decadal_2008}.

Bray-Curtis dissimilarities among samples revealed large differences in metabarcoding communities due to geographic Area ***(Hood Canal vs. San Juan Island; F = 1.6184; p $<$ 0.01). We therefore performed a constrained analysis of principal components (CAP) for samples within each Area, showing the differences among communities as a function of temperature, pH, and salinity (Fig. 2; for clarity, results for Hood Canal shown; full analysis in the Supporting Information).

@@

These communities were distinguished by changes in the relative abundances of a wide variety of taxa. In the Hood Canal for example, the community linked with colder water and higher-pH (community **2 in Fig. 2C) showed higher eDNA indexes of diatoms ***(\textit{Minutocellus polymorphus}), green algae  (\textit{Bathycoccus prasinos} and \textit{Micromonas pusilla}), and dinoflagellates like \textit{Karlodinium} sp relative to the lower-pH cluster at the same temperature. In that community (cluster 1) \textit{Emiliania huxleyi}, \textit{Ditylum brightwellii} and the copepod \textit{Paracalanus} sp. C are more prevalent.  The community common in the high temperature range (cluster 3 in Fig. 2C) shows high values of \textit{Phaeocystis globosa},\textit{Ditylum brightwellii} among other species. Cluster 3 (Fig. 2A,B) and similar planktonic communities occupy the spectrum of environmental conditions most likely to be encountered in near-future climate scenarios as temperature rises and pH falls. For example, we expect the Hood Canal in 2095 to have the conditions in which community 3 is the most likely community 66\% of the time, an increase of 11\% compared to 2017(see Supplemental Material). On the other hand, conditions that make cluster 2 the most likely community drop from 9\% to 1\% of the time between 2017 and 2095.

@@

Diatoms (Bacillariophyta) show the steepest richness decline under future conditions (Figure 3 B.2); the probability of occurrence decreases markedly for diatom taxa including \textit{Coscinodiscus} and \textit{Stephanodiscus}, both centric body forms. These declines in diatom richness were more accentuated at lower pH values and higher temperatures. Other taxa likely to find less-suitable habitat in the future include the dinoflagellate Gymnodiniaceae sp.

@@

 Changes in environmental conditions and associated shifts in planktonic communities will likely reshape ecosystems and food webs, although some environmental processes may be conserved even as the particular taxa change. A switch from a diatom-dominated ecosystem to one in which dinoflagellate blooms extend in space and time could provoke cascade effects \cite{xiao2018warming} including fish mortality, anoxia \cite{anderson2002harmful}, and carbon sinking dynamics \cite{bienfang1984sinking}. Beyond the phylum-specific patterns, the increase in suitable habitat for harmful algae species will alone be an engine for ecosystem change \cite{trainer2019pelagic, wells2018harmful}.

@@

The taxa surveyed here are a function of our metabarcoding PCR primers \cite{leray_new_2013} and reflect the current status of genetic databases, rather than a complete sampling of the planktonic community; we view these results as a cross-section of common taxa useful for understanding the biological effects of ocean conditions. Our observations are strong evidence of the kinds of changes likely in future marine communities, and they offer testable predictions about the magnitude and direction of effects on focal species.

@@

 We are grateful to the UW Center for Environmental Genomics and the Conservation Biology Molecular Genetics Laboratory at NOAA Fisheries (Montlake) for access to lab equipment and associated expertise and support. Thanks to Micah Horwith and the Washington State Department of Natural Resources for in-field support, to Terrie Klinger for feedback on an earlier draft, to Stephanie Moore for advice on phytoplankton generally, to Ana Ram\'{o}n-Laca for lab help, to Dan Drinan for field help, and to Alex Gagnon for access to carbonate-chemistry expertise and equipment.
